---
title: "Global Fertility and infant mortality"
author: "Chifundo Kanjala"
format:  
  dashboard:
    github: https://github.com/YOUR_URL
theme: lux
execute:
  echo: false
  warning: false
  message: false
---

```{r}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse,
               bslib, 
               here,
               bsicons, 
               shiny, 
               rgeoboundaries, 
               plotly, 
               gapminder,
               htmltools, 
               reactable, 
               janitor, 
               gganimate,
               gapminder, 
               rnaturalearth,
               countrycode)

```

```{r echo=FALSE, include=FALSE}

#import data

demo_indicators_raw <- 
  read_csv(here("data/demogr.csv"))

country_data <- 
  read_csv(here("data/CountriesGapminder.csv")) 

```


```{r echo=FALSE, message=FALSE}
country_iso_groups <- 
  country_data %>% 
  select(geo, name, four_regions) %>% 
  mutate(iso3_alpha_code = str_to_lower(geo))
```


```{r include=FALSE}
#initial cleaning
demo_indicators_cleaner1 <- demo_indicators_raw %>%
  janitor::row_to_names(row_number = 16) %>% 
    clean_names() %>% 
  rename(parent_code = na_4, year = na_5, pop_density = na_6, 
         total_pop = total_population_as_of_1_july_thousands, tfr = total_fertility_rate_live_births_per_woman, life_exp_0 = life_expectancy_at_birth_both_sexes_years,
         imr = infant_mortality_rate_infant_deaths_per_1_000_live_births) %>% 
  select(year,type, parent_code, iso3_alpha_code, region_subregion_country_or_area,pop_density, total_pop, tfr, life_exp_0,imr) %>% 
  mutate(total_pop = as.numeric(str_replace_all(total_pop, "\\s", "")), 
         iso3_alpha_code = str_to_lower(iso3_alpha_code))

demo_indicators_cleaner2 <- 
  demo_indicators_cleaner1 %>% 
  mutate(tfr = as.numeric(tfr),
         life_exp_0 = as.numeric(life_exp_0),
         imr = as.numeric(imr),
         parent_code = as.factor(parent_code))
```

```{r include=FALSE}
unique(demo_indicators_cleaner2$region_subregion_country_or_area)
setdiff(demo_indicators_cleaner2$iso3_alpha_code, country_iso_groups$iso3_alpha_code)
setdiff(country_iso_groups$iso3_alpha_code, demo_indicators_cleaner2$iso3_alpha_code)

countrycodes <- demo_indicators_cleaner2 %>% 
  select(region_subregion_country_or_area,iso3_alpha_code) %>% 
  filter(iso3_alpha_code %in% c("myt", "reu", "esh", "shn", "mac", "fro", 
                              "ggy", "imn", "jey", "gib", "vat", "xkx",
                              "aia", "abw", "bes", "vgb", "cym", "cuw",
                              "glp", "mtq", "msr", "pri", "blm", "maf",
                              "sxm", "tca", "vir", "flk", "guf", "bmu",
                              "grl", "spm", "ncl", "gum", "mnp", "asm", 
                              "cok", "pyf", "niu", "tkl", "wlf"))

unique(countrycodes$region_subregion_country_or_area)

```


```{r echo=FALSE, message=FALSE}

# 
demo_indicators_4regions <-
  inner_join(demo_indicators_cleaner2,country_iso_groups, by ="iso3_alpha_code")

```




```{r echo=FALSE}
library(ggplot2)
library(plotly)
library(dplyr)

# Convert total_pop to numeric
demo_indicators_simple_scatter <- demo_indicators_4regions %>%
  filter(type == "Country/Area") %>%
  mutate(total_pop = as.numeric(total_pop), four_regions = str_to_title(four_regions))  
  

# scatterplot
fert_imr_plot2 <- ggplot(data = demo_indicators_simple_scatter,
       mapping = aes(x = tfr,
                     y = imr,
                     frame = year,
                     color = four_regions,
                     size = total_pop,
                     text = paste("Country: ", region_subregion_country_or_area,
                                  "<br>Total Fertility Rate: ", tfr,
                                  "<br>Infant Mortality Rate: ", imr,
                                  "<br>Population, in 1000s: ", total_pop,
                                  "<br>Year: ", year,
                                  "<br>Region: ",four_regions))) +
  geom_point(alpha = 0.5) +
  scale_size(range = c(1, 20)) +
  #scale_x_log10()+
  theme_gray() +
  theme(legend.position = "none") +
  labs(title = "Infant Mortality Rate by Total Fertility Rate",
       y = "Infant Mortality Rate per 1000 live births",
       x = "Total Fertility Rate",
       color = "Region",
       size = "Total Population")

localeval <- Sys.setlocale(locale="C")
# Convert to plotly and define tooltips
fert_imr_plot2 <- ggplotly(fert_imr_plot2, tooltip = "text")

# Display the plot


```


```{r}
#fertmortplotly <- ggplotly(fert_imr_plot2, tooltip = "text")
```

```{r}
country_shapes <- rnaturalearth::ne_countries() %>% 
  mutate(adm0_a3 = str_to_lower(adm0_a3))


fert_imr_2021 <- demo_indicators_simple_scatter %>% 
  filter(year == 2021)


fert_map_2021 <-left_join(country_shapes,fert_imr_2021,by = c("adm0_a3" = "iso3_alpha_code")) %>% 
  mutate(tooltip_label = paste(
    admin,
    tfr, sep = ": ")) %>% 

  ggplot()+
  geom_sf(aes(fill = tfr, 
              text =tooltip_label))+
  theme_void()+
  theme(legend.position = "none")

fert_map_ggplotly <- ggplotly(fert_map_2021, tooltip = "text")
```


```{r echo=FALSE}
# highest fertility

highest_fert_country <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>% 
  arrange(-tfr) %>% 
  head(1) %>% 
  pull(region_subregion_country_or_area)

highest_fert_val <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>%
  arrange(-tfr) %>% 
  head(1) %>% 
  pull(tfr) %>% 
  round(1)

#lowest fertility
lowest_fert_country <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>% 
  arrange(tfr) %>% 
  head(1) %>% 
  pull(region_subregion_country_or_area)

lowest_fert_val <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>%
  arrange(tfr) %>% 
  head(1) %>% 
  pull(tfr) %>% 
  round(1)
```


```{r include=FALSE}
# highest infant mortality

highest_imr_country <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>%
  arrange(-imr) %>% 
  head(1) %>% 
  pull(region_subregion_country_or_area)

highest_imr_val <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>%
  arrange(-imr) %>% 
  head(1) %>% 
  pull(imr) %>% 
  round(1)

#lowest infant mortality
lowest_imr_country <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>%
  arrange(imr) %>% 
  head(1) %>% 
  pull(region_subregion_country_or_area)

lowest_imr_val <-
  demo_indicators_simple_scatter %>% 
  filter(year == 2021) %>%
  arrange(imr) %>% 
  head(1) %>% 
  pull(imr) %>% 
  round(1)
```


# HOME 

## Row 1 {height=25%}

```{r}
value_box(
  title = "Highest Total Fertility Rate, 2021",
  value = highest_fert_val,
  showcase = bsicons::bs_icon("person-plus"),
  theme = value_box_theme(bg = "#214773"),
  p(paste0("(", highest_fert_country,")"))
)
```


```{r}
value_box(
  title = "Lowest Total Fertility Rate, 2021",
  value = lowest_fert_val,
  showcase = bsicons::bs_icon("person-plus"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", lowest_fert_country,")"))
)
```




```{r}
value_box(
  title = "Highest Infant Mortality Rate, 2021",
  value = highest_imr_val,
  showcase = bsicons::bs_icon("clipboard-minus-fill"),
  theme = value_box_theme(bg = "#214773"),
  p(paste0("(", highest_imr_country,")"))
  
)
```


```{r}
value_box(
  title = "Lowest Infant Mortality Rate, 2021",
  value = lowest_imr_val,
  showcase = bsicons::bs_icon("clipboard-minus-fill"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", lowest_imr_country,")"))
  
)
```



## Row 2 {height=75%}

### {width=40%}


```{r title = "Total Fertility Rate by Infant Mortality Rate"}
fert_imr_plot2
```


### {width=60%}

```{r title = "Map of Countries by Total Fertility Rate, 2021"}
fert_map_ggplotly
```


# DOWNLOAD DATA

```{r echo=FALSE, message=FALSE}
library(htmltools)
demo_indicators_simple_scatter[] <- lapply(demo_indicators_simple_scatter, function(x) iconv(x, to = "UTF-8"))
htmltools::browsable(
  tagList(
      reactable(demo_indicators_simple_scatter,
                elementId = "demo_indicators_simple_scatter-table",
                searchable = T,
                filterable = T),
      
tags$button("Download as CSV",
            onclick = "Reactable.downloadDataCSV('demo_indicators_simple_scatter-table')")
  )
)


```



# ABOUT

These data come from the Gapminder website: https://www.gapminder.org/data/documentation/gd008/. They are estimates of the United Nations	Population Division	Department of Economic and Social Affairs, World Population Prospects 2022						
File GEN/01/REV1: Demographic indicators by region, subregion and country, annually for 1950-2100						
Estimates, 1950 - 2021						
POP/DB/WPP/Rev.2022/GEN/F01/Rev.1						
© July 2022 by United Nations, made available under a Creative Commons license CC BY 3.0 IGO: http://creativecommons.org/licenses/by/3.0/igo/						
Suggested citation: United Nations, Department of Economic and Social Affairs, Population Division (2022). World Population Prospects 2022, Online Edition.						
										
